---
title: "R Data Science Chapter 1"
author: "Shaun Jackson"
output: 
  rmarkdown::github_document:
    toc: true
    toc_depth: 2
---

The below notes are from Hadley Wickham's R for Data Science. This is a summary of my main takeaways from the sections that I thought were the most important

# Data visualisation

- The section of the book teaches us how to use ggplot2. In order to make sure all the required packages are available:

```{r message=F, warning=F}
library(tidyverse)
```

## Data we are working with

We are using the **mpg** dataframe that is found within ggplot2.
```{r}
mpg
```

- Note: mpg is a tibble. Tibbles are 

> "a modern reimagining of the data.frame." "Tibbles are data.frames that are lazy and surly: they do less" "and complain more." - from https://tibble.tidyverse.org 

## Plotting `mpg` by putting `displ` on the x-axis and `hwy` on the y-axis:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

In the above plot,

- `geom_point()` produces points on the graph.
- Define x and y axes by setting the x and y parameters within the aesthetic function `aes()` to `x = displ` and `y = hwy`.